Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2016-02-09 documents/ms/ms.tex
comments from Hogg

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to variations in the device.
There are also \emph{intra-pixel} sensitivity variations that can contribute\revise{(\citealt{subpixel})}.
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential \remove{(across the focal plane)} pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into light-curves \revise{\citep{handbook}}.
There are also \remove{probably} changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly other sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{ccd}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

\remove{We propose and advise dealing with}\revise{We propose to deal with} these variations in \Kepler\ light-curves by \emph{modeling} them.
In this context, a model is \remove{(at least)} a parameterized function that can predict data, given parameter settings,
  and an objective function that can be used to set or sample those parameters.
In some cases, the objective function can be given a probabilistic justification or interpretation, e.g., if it is constructed using a likelihood and a prior pdf.
The model can be a physical model (of the spacecraft PSF, pointing, temperature, and so on)
  or it can be a flexible, effective model that has no direct interpretation in terms of physical spacecraft parameters.
If done well, we would expect a physical model to do a better job,
  because it embodies more prior information,
  but it requires research and intuition about dominant effects. 
If this research and intuition is wrong or incomplete, a physical model may actually perform worse.
The model we propose here is in the non-physical, effective category.
The \Kepler\ community is familiar with these kinds of models;
  to our knowledge, \emph{all} successful light-curve ``de-trending'' methods
  are flexible, effective models.

@@

We describe the method and deliver all the relevant code in a public, open-source repository.
We also provide an interface to the \Kepler\ data that \remove{delivers improved ``\name\ photometry''}\revise{can be used
to produce ''\name\ photometry''}
  for every \Kepler\ target.

@@

Because of the train-and-test framework,  
  we expect the \name\ prediction $S_{n}$ to be a prediction of
  what the star light-curve would have under the influence of the spacecraft but with no exoplanet signal 
Based on this idea, 
  we can construct the ``\name\ flux'' 
  (where the inverted commas indicate that it is not a flux in the sense that stellar variability is not preserved) 
  to be the relative residual between \name\ prediction and SAP Flux $F_{n}$
\begin{eqnarray}
\delta_{n}&\equiv&\frac{F_{n} - S_{n}}{S_{n}}
\quad .
\end{eqnarray} 
In some sense, it is this relative residual--- the \name\ flux that contains the exoplanet transit signals we seek. 
That is, a transit creates a negative residual away from the prediction, 
  and the amount of relative residual is just the fraction of the light eclipsed by the planet (the transit depth).

\subsection{\revise{Tuning parameters}}
The \name\ has 4 \remove{hyper-parameters}\revise{tuning parameters}--- 
  the number of predictor stars Q (or number of predictor pixels P), 
  the number of autoregressive components S, 
  and the two regularization strengths $\lambda_{a}$ and $\lambda_{b}$.
To set these 4 \remove{hyper-parameters}\revise{tuning parameters}, 
  in principle we need to run something like a cross-validation on every single star to optimize performance.
But optimizing in this 4-dimensional space is expensive, 
  especially since in the \name\ we need to solve thousands of linear systems, 
  each of which has thousands of free parameters. 
Therefore, in this paper, 
  we just set a general set of ``default'' \remove{hyper-parameters}\revise{tuning parameters} ($P=4000$, \remove{$Q=3$}\revise{$S=3$}, $\lambda_a=1e5$, $\lambda_a=1e5$), 
  which works well in most of the stars without high variability. 
To show how this general set of \remove{hyper-parameters}\revise{tuning parameters} performs, 
  we present a few comparisons between optimized and default \remove{hyper-parameters}\revise{tuning parameters} 
  in Fig~\ref{hyperparameter}.
We can see that in Fig~\ref{hyperparameter}, 
  there is less variation in the \name\ flux 
  when we use optimized \remove{hyper-parameters}\revise{tuning parameters} (grey points in second panel, \revise{$Q=60$, $S=3$, $\lambda_a=1e7$, $\lambda_a=1e5$}) 
  than when we use the default one (red points in the bottom panel). 
In addition, the optimized flux has higher signal to noise ratio of the known transit than the default. 
However, the performance of default \remove{hyper-parameters}\revise{tuning parameters} is acceptable 
  and we will use this set of \remove{hyper-parameters}\revise{tuning parameters} throughout the rest of the paper 
  to show the general performance of \name. 
\revise{It is also worth to mention that the tuning parameters are optimized in a relatively coarse grid, which means the performance can still be improved when setting the tuning parameters more precisely.}
\revise{There are also parameters such as train-and-test window size, the CCD constraint on the predictor stars, the minimum distance of the predictor stars and the ranking algorithm to select predictor stars. In \name\ we set these parameters heuristically, but they could be open for discussion.}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{f3}
\end{center}
\caption{
  \label{hyperparameter} 
  Comparison between default and optimized \remove{hyper-parameters}\revise{tuning parameters}. 
  In the top panel, SAP flux is plotted in black, 
    \name\ prediction of default \remove{hyper-parameters}\revise{tuning parameters} is in thin red line, 
    \name\ prediction of optimized \remove{hyper-parameters}\revise{tuning parameters} is in bold grey line. 
  The middle panel shows the \name\ flux of the optimized \remove{hyper-parameters}\revise{tuning parameters}, 
    while \name\ flux of the default \remove{hyper-parameters}\revise{tuning parameters} is in the bottom panel, 
    the two vertical blue lines indicate the location of the transit signal, 
    and signal to noise ratio is calculated in both situation. 
  The optimized \remove{hyper-parameters}\revise{tuning parameters} performs better than the default one, 
    but the results from both situation are close.
  Here the signal strength is determined by mean depth of the transit signal,
    while the noise level is valued by 
    using the root-mean-square deviation of every points in a 6-day running window.}
\end{figure}

@@

\subsection{\name's effect on transit signals}
One important feature of \name\ is that we use the train-and-test framework to preserve 
the transit signal. To show how well train-and-test framework performs, we make the
following experiments on \Kepler\ data as shown in Fig~\ref{distortion}.
In the light-curve of \revise{KIC 9822284, simpy box model with different depth (from 100 ppm to 1\%) is injected}\remove{single pixel (1, 2) of KIC 9822284, a 10-hrs fake signal 
with amplitude 1.0005 is inserted. To insert the signal, we simply multiply the light-curve 
by the factor of 1.0005, which makes the little bump in the middle of the SAP light-curve 
(black points in the first panel).} With the distorted light-curve, both \name\ and "ordinary fitting" are applied the to the data. 
\revise{By "ordinary fitting", we mean fitting with all the data,  in comparison with the \name\, in which for each data point prediction is made without the data in the excluded region.}  
In the result, we can see that both methods fit the light-curve quite well outside the range of inserted signal. 
However, when examine how these two models preserve the signal that we care, the ''ordinary fitting'' fit out a large portion of the signal, while \name\ can preserve the whole 
original amplitude. Therefore,  by this simple experiment, we are able to show that the train-and-test framework used here is capable to keep the transit signal with a particular time scale. 
Also we want to emphasise that it is crucial for searching for
exoplanets, especially the small ones such as earth-like planets, which only have \remove{about tens of parts per million}\revise{$\backsim 100\ ppm$} signal in the light-curve. \revise{But there still exits some issue in the \name. 
The \name\ does not predict the light-curve very well around the box model, since in these regions the prediction is under the strong influence of the box model or "transit signal". 
Thus \name\ will introduce some distortion around the transit signal and the distortion increases with the depth of the signal. This issue will be discussed in detail in section 5.}

@@

To give a view on how our method performs, \name\ is applied on several stars with 
known transit signals but different variability and magnitudes as shown in Fig.~\ref{fluxes}. 
From the plots, we can see that \name\ works quite well on these stars, producing the \name\ fluxes 
with low variation and preserving the transit signals.
In addition, \name\ and PDC are compared based on 6 quiet stars 
(non-variable stars, first two rows of Fig.~\ref{fluxes}). 
The results illustrate that for these quiet stars our approach removes 
a major part of the variability present in the PDC light-curves, while preserving the transit signals. 
To provide a quantitative comparison, we ran \name\ on 1000 stars from the whole Kepler input catalog 
(500 chosen randomly from the whole list, and 500 random G-type sun-like stars), 
and estimate the Combined Differential Photometric Precision (CDPP,  \cite{cdpp1} )for \name\ and PDC.  
CDPP is an estimate of the relative precision in a time window, 
indicating the noise level seen by a transit signal with a given duration. 
\revise{In this paper, the CDPP is calculated with 12 bandpass filter using the equations 6-8 in section 3.3 of \citep{cdpp1}. And the CDPP of both PDC and \name\ is estimated with the same algorithm.} 
The duration is typically chosen to be 3, 6, or 12 hours. 
We use the 12-hours CDPP metric, since the transit duration of an earth-like planet is roughly 10 hours. 
Before CDPP is estimated, the variability of the stars are evaluated by median differential variability 
  (MDV, see \cite{basri2013} for details). 
Based on the variability, the 100 most variable stars in the list are removed, 
  since PDC is not designed to remove stellar variability, 
  and we want to compare \name\ and PDC on the close footing.
Fig.~\ref{cdpp} presents our CDPP comparison of \name\ and PDC, 
  \remove{showing that our method (\name) outperforms PDC.}\revise{4 quarters' (quarter 5, 6, 7, 8. Data of quarter 5 are from Kepler data release 21 and other data are from Kepler data release 24---the current version) data are used to show that \name\ has a consistent performance over a whole orbital period of the Kepler photometer (372 days). We also want to emphasis that both \name\ and PDC are presented here does not mean that it is a competition. Since PDC is designed to preserve stellar variability, it is not fair to put \name\ and PDC into competition. The CDPP comparison is made only to give a relative reference how \name\ can perform according to PDC, the Kepler "gold standard"}

>>>>

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to variations in the device.
There are also \emph{intra-pixel} sensitivity variations that can contribute\revise{(\citealt{subpixel})}.
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential \remove{(across the focal plane)} pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into light-curves \revise{\citep{handbook}}.
There are also \remove{probably} changes to the detector sensitivity with temperature and time,
  possibly illimination-history effects,
  and possibly other sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{ccd}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

\remove{We propose and advise dealing with}\revise{We propose to militate} these variations in \Kepler\ light-curves by \emph{modeling} them.
In this context, a model is \remove{(at least)} a parameterized function that can predict data, given parameter settings,
  and an objective function that can be used to set or sample those parameters.
In some cases, the objective function can be given a probabilistic justification or interpretation, e.g., if it is constructed using a likelihood and a prior pdf.
The model can be a physical model (of the spacecraft PSF, pointing, temperature, and so on)
  or it can be a flexible, effective model that has no direct interpretation in terms of physical spacecraft parameters.
If done well, we would expect a physical model to do a better job,
  because it embodies more prior information,
  but it requires research and intuition about dominant effects. 
If this research and intuition is wrong or incomplete, a physical model may actually perform worse.
The model we propose here is in the non-physical, effective category.
The \Kepler\ community is familiar with these kinds of models;
  to our knowledge, \emph{all} successful light-curve ``de-trending'' methods
  are flexible, effective models.

@@

We describe the method and deliver all the relevant code in a public, open-source repository.
We also provide an interface to the \Kepler\ data that \remove{delivers improved ``\name\ photometry''}\revise{can be used
to produce ``\name\ photometry''}
  for every \Kepler\ target.

@@

Because of the train-and-test framework,  
  we expect the \name\ prediction $S_{n}$ to be a prediction of
  what the star light-curve would have under the influence of the spacecraft but with no exoplanet signal. 
Based on this idea, 
  we can construct the ``\name\ flux'' 
  (where the inverted commas indicate that it is not a flux in the sense that stellar variability is not preserved) 
  to be the relative residual between \name\ prediction and SAP Flux $F_{n}$
\begin{eqnarray}
\delta_{n}&\equiv&\frac{F_{n} - S_{n}}{S_{n}}
\quad .
\end{eqnarray} 
In some sense, it is this relative residual\revise{---}the \name\ flux\revise{---}that contains the exoplanet transit signals we seek. 
That is, a transit creates a negative residual away from the prediction, 
  and the amount of relative residual is just the fraction of the light eclipsed by the planet (the transit depth).

\subsection{\revise{Tuning parameters}}
The \name\ has 4 \remove{hyper-parameters}\revise{tuning parameters}--- 
  the number of predictor stars Q (or number of predictor pixels P), 
  the number of autoregressive components S, 
  and the two regularization strengths $\lambda_{a}$ and $\lambda_{b}$.
To set these 4 \remove{hyper-parameters}\revise{tuning parameters}, 
  in principle we need to run something like a cross-validation on every single star to optimize performance.
But optimizing in this 4-dimensional space is expensive, 
  especially since in the \name\ we need to solve thousands of linear systems, 
  each of which has thousands of free parameters. 
Therefore, in this paper, 
  we just set a general set of ``default'' \remove{hyper-parameters}\revise{tuning parameters} ($P=4000$, \remove{$Q=3$}\revise{$S=3$}, $\lambda_a=1e5$, $\lambda_a=1e5$), 
  which works well in most of the stars without high variability. 
To show how this general set of \remove{hyper-parameters}\revise{tuning parameters} performs well, 
  we present a few comparisons between optimized and default \remove{hyper-parameters}\revise{tuning parameters} 
  in Fig~\ref{hyperparameter}.
We can see that in Fig~\ref{hyperparameter}, 
  there is less variation in the \name\ flux 
  when we use optimized \remove{hyper-parameters}\revise{tuning parameters} (grey points in second panel, \revise{$Q=60$, $S=3$, $\lambda_a=1e7$, $\lambda_a=1e5$}) 
  than when we use the default \remove{one}\revise{values} (red points in the bottom panel). 
In addition, the optimized flux has higher signal-to-noise ratio \remove{of}\revise{on} the known transit than the default. 
However, the performance of default \remove{hyper-parameters}\revise{tuning parameters} is acceptable 
  and we will use this set of \remove{hyper-parameters}\revise{tuning parameters} throughout the rest of the paper 
  to show the general performance of \name. 
\revise{It is also worth mentioning that the tuning parameters are optimized only on a relatively coarse grid, which means the performance can still be improved when setting the tuning parameters more precisely.}
\revise{There are also parameters such as train-and-test window size, the CCD constraint on the predictor stars, the minimum distance of the predictor stars and the ranking algorithm to select predictor stars. In \name\ we set these parameters heuristically, but they could be open for discussion.}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{f3}
\end{center}
\caption{
  \label{hyperparameter}
  Comparison between default and optimized \remove{hyper-parameters}\revise{tuning parameters}. 
  In the top panel, SAP flux is plotted in black, 
    \name\ prediction of default \remove{hyper-parameters}\revise{tuning parameters} is in thin red line, 
    \name\ prediction of optimized \remove{hyper-parameters}\revise{tuning parameters} is in bold grey line. 
  The middle panel shows the \name\ flux of the optimized \remove{hyper-parameters}\revise{tuning parameters}, 
    while \name\ flux of the default \remove{hyper-parameters}\revise{tuning parameters} is in the bottom panel, 
    the two vertical blue lines indicate the location of the \revise{injected} transit signal, 
    and the signal to noise ratio is calculated in both situations. 
  The optimized \remove{hyper-parameters}\revise{tuning parameters} perform better than the default\remove{ one}, 
    but the results from both situations are \remove{close}\revise{similar}.
  Here the signal strength is determined by the mean depth of the transit signal,
    while the noise level is \remove{valued by using}\revise{estimated with} the root-mean-square deviation of the points in a 6-day running window.}
\end{figure}

@@

\subsection{Effect on transit signals}
One important feature of \name\ is that we use the train-and-test framework to preserve 
the transit signal. To show how well the train-and-test framework performs, we \remove{make}\revise{perform} the
following experiments on \Kepler\ data shown in Fig~\ref{distortion}:
In the light-curve of \revise{KIC 9822284, simple rectangular box model with different depth (from 100 ppm to 1\%) is injected}\remove{single pixel (1, 2) of KIC 9822284, a 10-hrs fake signal 
with amplitude 1.0005 is inserted. To insert the signal, we simply multiply the light-curve 
by the factor of 1.0005, which makes the little bump in the middle of the SAP light-curve 
(black points in the first panel).} With the distorted light-curve, both \name\ and ``ordinary fitting" are applied to the data. 
\revise{By ``ordinary fitting", we mean fitting with all time points,  in comparison with the \name\, in which for each data point the prediction is made without the data in the excluded region.}  
\remove{In the result, we can see that} Both methods fit the light-curve quite well outside the range of the inserted signal. 
However, \remove{when examine how these two models preserve the signal that we care,} the ``ordinary fitting'' fits out a large portion of the signal, while the \name\ \remove{can} preserves the \remove{whole} original amplitude. 
\remove{Therefore,  by} This simple experiment\remove{, we are able to} shows that the train-and-test framework \remove{used here} is capable \remove{to keep}\revise{of preserving} the transit signal \remove{with a particular time scale}. 
\remove{Also we want to emphasise that it} This is crucial for searching for
exoplanets, especially \remove{the small ones such as} Earth-like planets, which only have \remove{about tens of parts per million}\revise{$\backsim$ 100\ ppm} signals \remove{in the light-curve}. \revise{But there are issues: 
The \name\ does not predict the light-curve very well around the box model, since in these regions the prediction is under the strong influence of the transit signal. 
Thus the \name\ will introduce some distortion around the transit signal, and \revise{we expect that} the distortion increases with the depth of the signal. This issue will be discussed in detail in Section 5.}

@@

\remove{To give a view on how our method performs,}\revise{The} \name\ is applied on several stars with known transit signals but different variability and magnitudes, as shown in Fig.~\ref{fluxes}. 
\remove{From the plots, we can see that}\revise{The} \name\ works \remove{quite} well on these stars, producing \remove{the} \name\ fluxes 
with low variation and preserving the transit signals.
In addition, \name\ and PDC are compared \remove{based}\revise{for} on 6 quiet stars 
(non-variable stars, first two rows of Fig.~\ref{fluxes}). 
The results illustrate that for these quiet stars our approach removes 
\remove{a major part}\revise{the majority} of the variability present in the PDC light-curves, while preserving the transit signals. 
To provide a quantitative comparison, we ran \name\ on 1000 stars from the whole Kepler input catalog 
(500 chosen randomly from the whole list, and 500 random G-type sun-like stars), 
and estimated the Combined Differential Photometric Precision (CDPP,  \citealt{cdpp1} )for the \name\ and PDC.  
CDPP is an estimate of the relative precision in a time window, 
indicating the noise level seen by a transit signal with a given duration. 
\revise{In this paper, the CDPP is calculated by using the equations 6-8 in section 3.3 of \citep{cdpp1}. The CDPP of both PDC and \name\ is estimated with the same algorithm.} 
The duration is typically chosen to be 3, 6, or 12 hours. 
We use the 12-hour CDPP metric, since the transit duration of an earth-like planet is roughly 10 hours. 
Before the CDPP is estimated, the variability of the stars \remove{are}\revise{is} evaluated by median differential variability 
  (MDV, see \citealt{basri2013} for details). 
Based on the variability, the 100 most variable stars in the list are removed, 
  since PDC is not designed to remove stellar variability, 
  and we want to compare \name\ and PDC on the close footing.
Fig.~\ref{cdpp} presents our CDPP comparison of \name\ and PDC, 
  \remove{showing that our method (\name) outperforms PDC.}\revise{4 quarters' (quarter 5, 6, 7, 8. Data of quarter 5 are from Kepler data release 21 and other data are from Kepler data release 24---the current version) data are used to show that \name\ has a consistent performance over a whole orbital period of the Kepler photometer (372 days). We also want to emphasis that both \name\ and PDC are presented here does not mean that it is a competition. Since PDC is designed to preserve stellar variability, it is not fair to put \name\ and PDC into competition. The CDPP comparison is made only to give a relative reference how \name\ can perform according to PDC, the Kepler "gold standard"}
